Classless Java

### 2015-11-05 paper/sections/Formalization.tex
Translation refactoring.
Refactor more.
Translation function fixed

\begin{figure}[h]
\begin{grammar}
\production{
\e
}{
  \x\mid\MCall\e\m\es\mid\MCall{\C}\m\es\mid\MCall{\C\QM{.super}}\m\es\mid\obj
  }{expressions}\\
\production{
\obj
}{
\QM{new}\ \C\oR\cR\oC\fields\ 
\mh_1\oC\QM{return}\ \e_1\QM{;}\!\cC
\ldots
\mh_n\oC\QM{return}\ \e_n\QM{;}\!\cC
\cC
  }{object creation}\\
\production{\field}{\T\ \f \QM= \x\QM;}{field declaration}\\
\production{
\metaVar{I}
}{
 \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_k \oC\method_1\ldots\method_n\cC
  }{interface declaration}\\
\production{
\method
}{
 \QM{static}\ \mh\ \oC\QM{return}\ \e\QM{;}\!\cC
\mid
\QM{default}\ \mh\oC\QM{return}\ \e\QM{;}\cC
\mid
\mh\QM{;}\quad
  }{method declaration}\\
\production{
\mh
}{
 \T_0\ \m\ \oR\T_1\ \x_1\ldots\T_n\ \x_n\cR
  }{method header}\\
\production{
\ann
}{
  \mixinAnn|\emptyset
  }{annotations}\\
\production{\Gamma}{
\x_1{:}\C_1\ldots\x_n{:}\C_n
}{environment}
\end{grammar}
\caption{Grammar of ClassLess Java}
\label{Grammar}
\end{figure}

@@

@@

\yanlin{for \textbf{shadow} conflicted mh;, one missing case is when there are
  two conflicting default methods with same signature?? }

@@

\[ [\![\mixinAnn\ \QM{interface}\ C_0\ \QM{extends}\ \Cs \oC \methods \cC ]\!] = \]
\[ \emptyset\ \QM{interface}\ C_0\ \QM{extends}\ \Cs \oC \methods;\ \ofMethod(C_0);\ \otherMethod(C_0) \cC \]

\haoyuan{Not $\{ \methods; ... \}$, actually. Should avoid duplication.}

\textbf{THEOREM. } If $I \longmapsto I'$ and $I$ OK, then $I'$ OK.

@@

(4) If $meth'=\none$, we need to figure out
\[\textsf{mostSpecific}(\overline{meth})\]
and it should be the one that ``overrides'' all the others in $\overline{meth}$. It means we should not only deal with the return types of methods, but also look into the subtyping relation of interfaces. But for abstract methods, only return types are taken into consideration.
}

\begin{figure}[tbp]
\centering
\includegraphics[width=5in]{table.png}
\caption{Table.}\label{table_png}
\end{figure}
\end{comment}

@@

\subsubsection{Auxiliary function: $\ofMethod(C_0)$, $\otherMethod(C_0)$}

\begin{equation*}
\begin{array}{ll}
\ofMethod(C_0) = \QM{static}\ C_0\ \QM{of} \oR T_1\ m_1\QM,\ldots T_k\ m_k\cR\ \QM{\{} &  \\
\ \ \ \ \QM{return new}\ C_0 \oR\cR\ \QM{\{}  & \mif\ \resolve(C_0) = \methods \\
\ \ \ \ \ \ \ \ T_1\ \_m_1 = m_1\QM;\ldots T_k\ \_m_k = m_k\QM; & \forall \method_i \in \methods, \method_i = T_i\ m_i\oR\cR\QM; \\
\ \ \ \ \ \ \ \ T_i\ m_i\oR\cR\ \QM{\{return}\ \_m_i\QM{;\}}\ \ldots & T_i\ m_i\oR\cR \text{ included for all } i \\
\ \ \ \ \ \ \ \ C_0\ \QM{with}m_i\oR T_i\ x_i\cR\ \QM{\{} & C_0\ \QM{with}m_i\oR T_i\ x_i\cR \text{ included if }\with(\method_i, C_0) \text{ defined} \\
\ \ \ \ \ \ \ \ \ \ \ \ \QM{return this.of}\oR m_1\oR\cR\QM,\ldots x_i\QM,\ldots m_k\oR\cR\cR\QM{;} & \\
\ \ \ \ \ \ \ \ \QM{\}}\ \ldots & \\
\ \ \ \ \ \ \ \ C_0\ \QM{clone}\oR\cR\ \QM{\{} & C_0\ \QM{clone}\oR\cR \text{ included if }\clone(C_0) \text{ defined} \\
\ \ \ \ \ \ \ \ \ \ \ \ \QM{return this.of}\oR m_1\oR\cR\QM,\ldots m_k\oR\cR \cR\QM; & \\
\ \ \ \ \ \ \ \ \QM{\}} & \\
\ \ \ \ \QM{\};} & \\
\QM{\}} & \\
\otherMethod(C_0) = \overline{\with(\method, C_0)}; \clone(C_0) \hspace{0.1in} & \mif\ \textsf{resolve}(C_0) = \methods \\
\resolve(C_0) = \{\mBody(m, C_0)\ |\ \mBody(m, C_0)\ \isField\} & \forall m\in\dom(C_0), \text{one case is satisfied:} \\
& (1)\ \mBody(m, C_0)\ \isField \\
& (2)\ \mBody(m, C_0)\ \isWith \\
& (3)\ \mBody(m, C_0)\ \isClone \\
& (4)\ \mBody(m, C_0)\ \isDefault \\
\mBody(m, C_0)\ \isField & \mif\ \mBody(m, C_0) = T\ m\oR\cR\QM;\ m\ne\QM{clone} \\
\mBody(m, C_0)\ \isWith & \mif\ \mh^{\mBody(m, C_0)} = T'\ \QM{with}m' \oR T\ x\cR, C_0 <: T' \\
& \mand\ \mBody(m', C_0) = T\ m'\oR\cR\QM;\ m'\ne\QM{clone} \\
\mBody(m, C_0)\ \isClone & \mif\ \mh^{\mBody(m, C_0)} = T\ \QM{clone}\oR\cR, C_0 <: T \\
\mBody(m, C_0)\ \isDefault & \mif\ \mBody(m, C_0) = \QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \\
\with(\method, C_0) = C_0\ \QM{with}m\oR T\ x\cR\QM; & \mif\ \method = T\ m\oR\cR\QM;\ \mBody(\QM{with}m, C_0)\ \isWith \\
\clone(C_0) = C_0\ \QM{clone}\oR\cR\QM; & \mif\ \mBody(\clone, C_0)\ \isClone \\
\end{array}
\end{equation*}

>>>>

\begin{figure}[h]
\begin{grammar}
\production{
\e
}{
  \x\mid\MCall\e\m\es\mid\MCall{\C}\m\es\mid\MCall{\C\QM{.super}}\m\es\mid\obj
  }{expressions}\\
\production{
\obj
}{
\QM{new}\ \C\oR\cR\oC\fields\
\mh_1\oC\QM{return}\ \e_1\QM{;}\!\cC
\ldots
\mh_n\oC\QM{return}\ \e_n\QM{;}\!\cC
\cC
  }{object creation}\\
\production{\field}{\T\ \f \QM= \x\QM;}{field declaration}\\
\production{
\II
}{
 \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_k \oC\method_1\ldots\method_n\cC
  }{interface declaration}\\
\production{
\method
}{
 \QM{static}\ \mh\ \oC\QM{return}\ \e\QM{;}\!\cC
\mid
\QM{default}\ \mh\oC\QM{return}\ \e\QM{;}\cC
\mid
\mh\QM{;}\quad
  }{method declaration}\\
\production{
\mh
}{
 \T_0\ \m\ \oR\T_1\ \x_1\ldots\T_n\ \x_n\cR
  }{method header}\\
\production{
\ann
}{
  \mixinAnn|\emptyset
  }{annotations}\\
\production{\Gamma}{
\x_1{:}\C_1\ldots\x_n{:}\C_n
}{environment}
\end{grammar}
\caption{Grammar of ClassLess Java}
\label{Grammar}
\end{figure}

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=5in]{table.png}
\caption{Table.}\label{table_png}
\end{figure}

@@

@@

$[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs \oC \methods \cC ]\!] =$\\*
${}_{}$\quad$\emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC 
\methods\ 
\ofMethod(\C_0) \ 
\otherMethod(\C_0)\ \cC
$\\*\tab
where  $\valid(\C_0)$
%\item $\overline{T\ m\oR\cR\QM;} = \fieldsFunc(C_0)$
%\item $p_i = \isWith(\mh^{\mBody(\QM{with#}m_i, C_0)}, C_0)$
%\item $q = \isClone(\mh^{\mBody(\clone, C_0)}, C_0)$
%\end{itemize}

@@

(4) If $meth'=\none$, we need to figure out
\[\textsf{mostSpecific}(\overline{meth})\]
and it should be the one that ``overrides'' all the others in $\overline{meth}$. It means we should not only deal with the return types of methods, but also look into the subtyping relation of interfaces. But for abstract methods, only return types are taken into consideration.
}
\end{comment}

@@

\subsubsection{Auxiliary function: $\valid, \ofMethod$, $\otherMethod$}${}_{}$\\*

\noindent$\valid(C_0)$  holds if $\forall \m\in\dom(\C_0),$ if $\mh\QM; = \mBody(\m, \C_0),$ one case is satisfied:
$\isField(\method)$,
$\isWith(\method, \C_0)$
or $\isClone(\method, \C_0)$.
%or $\isImplemented(\method)$.

We write $\QM{with#}\m$ to append $\m$ to \QM{with}, following the camelCase rule, so the first letter of
$\m$ must be lowercase and is turned in uppercase upon merging.
For example \QM{with#foo}=\QM{withFoo}.
Special names $\specialName(\m)$ are \QM{clone},\QM{with} and all the identifiers of form $\QM{with#}\m$.

We formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory. To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that all the methods do not start with underscore, and we prefix method names with underscore to obtain valid field/parameter names.\\*
\noindent$\begin{array}{l}
\ofMethod(\C_0) = \
 \QM{static}\ C_0\ \QM{of} \oR \C_1\ \m_1\QM,\ldots \C_n\ \m_n\cR\
\QM{\{}
\QM{return new}\ C_0 \oR\cR\ \QM{\{} \\
\tab\tab\tab\tab\tab\tab\tab C_1\ \QM_\m_1 = \m_1\QM;\ldots \C_n\ \QM_\m_n = \m_n\QM; \\
\tab\tab\tab\tab\tab\tab\tab
\C_1\ \m_1\oR\cR\ \QM{\{return _}\m_1\QM{;\}}\ \ldots 
\C_n\ \m_n\oR\cR\ \QM{\{return _}\m_n\QM{;\}}\\
\tab\tab\tab\tab\tab\tab\tab\withMethod(\m_1,\C_0,\es_1)\ldots\withMethod(\m_n,\C_0,\es_n)\\
\tab\tab\tab\tab\tab\tab\tab\cloneMethod(\C_0,\es)\\
\tab\tab\tab\tab\tab\tab\QM{\};\}} \\
\end{array}$
\\*
with $\fieldsFunc(\C_0)=\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();}$,\\*
$\es_i=\QM{this.}\m_1\QM{(),}\ldots \QM{this.}\m_{1-i}\QM{(),}\QM{_val,}\QM{this.}\m_{i+1}\QM{(),}\ldots \QM{this.}\m_n\oR\cR$\\*
and $\es=\QM{this.}\m_1\QM{(),}\ldots\ldots \QM{this.}\m_n\oR\cR$\\*

\noindent{}Other auxiliary functions:\\*
\noindent$\begin{array}{ll}
\withMethod:&\withMethod(\C,\m,\C_0,\es)=\method=
\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\ \QM{\{}
\QM{return} \C_0\QM{.of(}\es\QM{);\}} \\
&\mbox{iff }\method<:\mBody(\QM{with#}\m,\C_0)\\
&\withMethod(\C,\m,\C_0,\es)=\emptyset\mbox{ otherwise}\\
\cloneMethod:&\cloneMethod(\C_0,\es)=\method=
\C_0\ \QM{clone()\{return}\ \C_0\QM{.of(}\es\QM{);\}} \\
&\mbox{iff }\method<:\mBody(\QM{clone},\C_0)\\
&\cloneMethod(\C_0,\es)=\emptyset\mbox{ otherwise}\\

\otherMethod:& \C_0\ \QM{with#}\m\oR \C_i\ \x_i\cR\QM;\in
\otherMethod(\C_0) 
\\&
 \mbox{iff }
\C\ \m\QM{();}\in \fieldsFunc(\C_0)\mbox{ and } \QM{with#}\m\notin\dom(\C_0)\\
&\C_0\ \QM{clone}\oR\cR\QM;\in
\otherMethod(\C_0)  \mbox{ iff } \isClone( \mBody(\m, \C_0))\\

%other methos was wrong: we need the with to be NOT there to generate it.
%no use isWith, any method with that name must be absent

\fieldsFunc:&\method\in\fieldsFunc(\C_0) \mbox{ iff }
\method\in \dom(\C_0)\ \mand\ \isField(\method)
\\
\isField:&\isField(\C\ \m\oR\cR\QM;)\tab \mif\ \mnot\ \specialName(\m)\\
\isWith:&\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)
\tab \mif\ \C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;\\
& \mand\ \mnot\ \specialName(\m)\\
\isClone:&\isClone(\C\ \QM{clone}\oR\cR\QM;, \C_0)\tab \mif\ \C_0 <: \C \\
%\isImplemented:&\isImplemented(\method) \tab\mbox{iff }\method\mbox{ not of form }\mh\QM;
%\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
%&\isImplemented(\QM{static}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
\end{array}$

\subsubsection{Results}
\textbf{THEOREM. } 
For a given $\II_0\ldots\II_n$ interface table such that
$\forall\II\in\II_0\ldots\II_n, \II$ OK, then in the interface table
$[\![\II_0]\!]\II_1\ldots\II_n$
$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n, \II$ OK.
